---
title: "Reproducable Research- Assignment 1"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Loading and processing the data

```{r echo=TRUE}
getwd()
library(knitr)
library(dplyr)
library(ggplot2)
rr <- read.csv(file="activity.csv", header=TRUE)

# Remove NAs from data

rr_clean <- rr[ with (rr, { !(is.na(rr$steps)) } ), ]
```

## 2. Mean total number of steps taken per day (ignoring NA values)

```{r echo=TRUE}
steps_day <- group_by(rr_clean, date)

steps_per_day <- summarise(steps_day, total = sum(steps))

# Histogram for total steps in each day

hist(steps_per_day$total, main="Histogram for total number of steps each day", xlab="Total steps in one day")

```

# Mean, median of total number of steps taken
```{r echo=TRUE}
summary(steps_per_day)

```

## 3. Average daily activity pattern
## Time series plot

```{r echo=TRUE}
steps_interval <- aggregate(steps ~ interval, rr_clean, mean)
 
plot(steps_interval$interval, steps_interval$steps, type='l', main="Average daily activity pattern", xlab="Interval", ylab="Average number of steps")

```

## Interval with maximum number of steps
```{r echo=TRUE}
max_steps_row <- which.max(steps_interval$steps)
 
steps_interval[max_steps_row, ]
```

## 4.Dealing with missing values

## Total number of missing values
```{r echo=TRUE}
sum(is.na(rr))
```

## Filling the missing values with the mean of 5-minute interval
```{r echo=TRUE}
data_impute <- rr
for (i in 1:nrow(data_impute)) {
     if (is.na(data_impute$steps[i])) {
         interval_value <- data_impute$interval[i]
         steps_value <- steps_interval[
             steps_interval$interval == interval_value,]
         data_impute$steps[i] <- steps_value$steps
     }
 }
 
 
 imputed_steps <- aggregate(steps ~ date, data_impute, sum)

```

## Histogram for total number of steps taken each day
```{r echo=TRUE}
hist(imputed_steps$steps, main="Histogram of total number of steps per day (imputed)", 
     xlab="Total number of steps for each day")
```

##Mean, median of steps taken per day
###For imputed data
```{r echo=TRUE}
mean(imputed_steps$steps)
median(imputed_steps$steps)

```
###Data without NA values
```{r echo=TRUE}
mean(steps_per_day$total)
median(steps_per_day$total)

```
### So the mean is same but different median

## 5. Difference in activity patterns between weekdays and weekends

```{r echo=TRUE}
data_impute['type_of_day'] <- weekdays(as.Date(data_impute$date))
 data_impute$type_of_day[data_impute$type_of_day  %in% c('Saturday','Sunday') ] <- "weekend"
 data_impute$type_of_day[data_impute$type_of_day != "weekend"] <- "weekday"
 


 # convert type_of_day from character to factor
 data_impute$type_of_day <- as.factor(data_impute$type_of_day)
 
 # calculate average steps by interval across all days
 imputed_steps <- aggregate(steps ~ interval + type_of_day, data_impute, mean)

```

##Panel plot containing time series
```{r echo=TRUE}
# creating panel plot
 qplot(interval, 
       steps, 
       data = imputed_steps, 
       type = 'l', 
       geom=c("line"),
       xlab = "Interval", 
       ylab = "Number of steps", 
       main = "") +
    facet_wrap(~ type_of_day, ncol = 1)
```

